```{r label = "setup", include = FALSE}
library(nycflights13)
library(ggplot2)
library(dplyr)
library(qwraps2)
options(qwraps2_markup = "markdown")
knitr::opts_chunk$set(echo = FALSE, results = "hide")
```

# NYC Flights in 2013

We are interested in looking at the flights from departing NYC in 2013 and
arriving in Denver Colorado.  For this work we will use the data provided in the
R package [nycflights13](https://cran.r-project.org/package=nycflights13).
```{r, include = FALSE}
# There are several data sets we could use
data(package = "nycflights13")$results
```

## Fights To Denver
```{r, include = FALSE}
flights_to_den <-
  flights %>%
  dplyr::filter(dest == "DEN") %>%
  left_join(airlines, by = "carrier")

dplyr::filter(airports, faa %in% unique(flights_to_den$origin))
```

There were `r frmt(nrow(flights_to_den))` flights from NYC to Denver during
2013. `r n_perc(flights_to_den$origin == "EWR")` of the flights originated from
`r airports[airports$faa == "EWR", ][["name"]]`,
`r n_perc(flights_to_den$origin == "JFK")` originated from
`r airports[airports$faa == "JFK", ][["name"]]`, and the last
`r n_perc(flights_to_den$origin == "LGA")` of flights originated from
`r airports[airports$faa == "LGA", ][["name"]]`.

A summary of the flights are provided in the table below:

```{r, results = "asis"}
# with(flights_to_den, table(carrier, name))
smmry <-
  list("Departure Delay (minutes)" =
       list("minimum"      = ~ qwraps2::frmt(min(na.omit(.data[["dep_delay"]]))),
            "median (IQR)" = ~ qwraps2::median_iqr(na.omit(.data[["dep_delay"]])),
            "mean (sd)"    = ~ qwraps2::mean_sd(na.omit(.data[["dep_delay"]])),
            "maximum"      = ~ qwraps2::frmt(max(na.omit(.data[["dep_delay"]]))),
            "Unknown (%)"  = ~ qwraps2::n_perc(is.na(.data[["dep_delay"]]), digits = 0, show_symbol = FALSE, show_denom = "always")
            ),
       "Arrival Delay (minutes)" =
       list("minimum"      = ~ qwraps2::frmt(min(na.omit(.data[["arr_delay"]]))),
            "median (IQR)" = ~ qwraps2::median_iqr(na.omit(.data[["arr_delay"]])),
            "mean (sd)"    = ~ qwraps2::mean_sd(na.omit(.data[["arr_delay"]])),
            "maximum"      = ~ qwraps2::frmt(max(na.omit(.data[["arr_delay"]]))),
            "Unknown (%)"  = ~ qwraps2::n_perc(is.na(.data[["arr_delay"]]), digits = 0, show_symbol = FALSE, show_denom = "always")
            ),
       "Carrier" =
       list("JetBlue Airways"        = ~ qwraps2::n_perc(.data[["carrier"]] == "B6", digits = 0, show_symbol = FALSE),
            "Delta Air Lines Inc."   = ~ qwraps2::n_perc(.data[["carrier"]] == "DL", digits = 0, show_symbol = FALSE),
            "Frontier Airlines Inc." = ~ qwraps2::n_perc(.data[["carrier"]] == "F9", digits = 0, show_symbol = FALSE),
            "United Air Lines Inc."  = ~ qwraps2::n_perc(.data[["carrier"]] == "UA", digits = 0, show_symbol = FALSE),
            "Southwest Airlines CO." = ~ qwraps2::n_perc(.data[["carrier"]] == "WN", digits = 0, show_symbol = FALSE)
            )
  )

cbind(
      summary_table(flights_to_den, smmry),
      summary_table(dplyr::group_by(flights_to_den, origin), smmry)
     )
```

A graphic for the flight times...

```{r, warning = FALSE, fig.width = 15, fig.height = 7}
ggplot(flights_to_den) +
  aes(x = time_hour, y = air_time, color = carrier) +
  geom_point() +
  geom_line() +
  facet_grid( ~ origin)
```

