---
title: "Death Maps"
author: "Peter DeWitt"
package: drug.eg
output:
 BiocStyle::html_document:
    toc_float: true
vignette: >
 %\VignetteIndexEntry{Death Maps}
 %\VignetteEngine{knitr::rmarkdown}
 %\VignetteEncoding{UTF-8}
---


```{r label = "setup", include = FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```


This is a proof of concept vignette.  This document has been written to show
that knitr::spin can be used to build vignettes.

# Deaths in the United States

This is a brief exploration of the top 10 causes of death in the United
States over the past several years.


```{r }
library(drug.eg)
library(ggplot2)
library(maps)
library(mapproj)
library(dplyr)

data(all_cause_states, by_cause_states)

# map data
us_states <- map_data("state") %>% dplyr::as_tibble(.)
```


# All Cause Deaths


```{r }
dplyr::left_join(us_states,
                 {
                 all_cause_states %>%
                   dplyr::mutate(State = tolower(State))
                 },
                 by = c("region" = "State")) %>% 
ggplot(.) +
  theme_bw() +
  aes(x = long, y = lat, group = group, fill = `Age-adjusted Death Rate`) +
  geom_polygon() +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  theme(
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.position = "bottom",
        legend.key.width = grid::unit(0.5, "in")) +
  facet_wrap( ~ Year)
```


# By Cause Deaths

Cancer and Heart Disease are the two most common causes of death.

```{r fig.width = 12, fig.height = 6}
by_cause_us %>%
  dplyr::group_by(Year) %>%
  dplyr::mutate(Rank = rank(-`Age-adjusted Death Rate`)) %>%
  dplyr::filter(Rank <= 3) %>%
ggplot(.) +
  theme_bw() +
  geom_bar(aes(x = Year, y = `Age-adjusted Death Rate`, fill = `Cause Name`),
           position = position_dodge(),
           stat = "identity") +
  theme(legend.position = "bottom")
```


## Heart Disease

The graphics below show the map of the United states and the age adjusted
death rate for cancer and heart disease over the years of interest.

```{r }
dplyr::left_join(us_states,
                 {
                   by_cause_states %>%
                     dplyr::filter(`Cause Name` %in% c("Cancer", "Heart disease")) %>%
                     dplyr::mutate(State = tolower(State))
                 },
                 by = c("region" = "State")) %>%
ggplot(.) +
theme_bw() +
aes(x = long, y = lat, group = group, fill = `Age-adjusted Death Rate`) +
geom_polygon() +
coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
theme(
      axis.title = element_blank(),
      axis.text = element_blank(),
      axis.ticks = element_blank(),
      legend.position = "bottom",
      legend.key.width = grid::unit(0.5, "in")) +
facet_wrap( ~ Year + `Cause Name`)
```

